---
title: "HS students"
author: "Anthony Gutierrez"
date: "5/14/2021"
output:
  html_document:
    rmdformats::readthedown
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# Load Libraries 

```{r}
library(tidyverse)
library(here)
library(tidytuesdayR) 
library(patchwork)
```

# Load Data

```{r}
tuesdata3 <- tt_load("2021-02-02") #Load in TidyT data

hbcu_all <- tuesdata3$hbcu_all #look at all college graduations 

```

# Introduction
The goal of this Tidy Tuesday is to plot the number of college graduates by gender and type of college that they have attended.


# Manipulate data

## Manipulate data to plot graduation by gender
```{r}
hbcu_by_gender <- hbcu_all %>%
  # We only need year and gender columns
  select(Year, Males, Females) %>%
  # Convert to tidy format, collapsing male/female into one descriptor field
  pivot_longer(Males:Females,
               names_to = "gender",
               values_to = "students") %>%
  # Convert from plural to singular for cleaner data
  # "s%" specifies an s character at the end of a string
  # ("$" is end of string in regular expressions)
  mutate(gender = str_remove(gender, "s$"))

```

## Plot graduation numbers by gender
```{r}
gender_grad_num <- hbcu_by_gender %>% #create ggplot for graduating gender statistics
  ggplot(aes(x = Year,
             y = students,
             group= gender, #group by gender
             color= gender))+ #plot number of grads by year
  geom_line()+ #create line plot
  labs(title = "Number of graduating students by gender",
       y = "Graduating Students")+ #add descriptive titles to figure and legends
  theme_classic()
  
```



## Manipulate data to plot by college type

```{r}
hbcu_by_school <- hbcu_all %>%
  select(1,2,8,9,11,12) %>% # We only need year and college type columns
  pivot_longer(3:6, 
               names_to = "College",
               values_to = "students") #collapsing college type into one descriptor field


```

```{r}
grad_rate <- hbcu_all %>%
  select(1,2,5,6) %>% # We only need year and college type columns
  pivot_longer(3:4,
               names_to = "College",
               values_to = "students")%>%  #collapsing college type into one descriptor field
  mutate(rate = 100*students/`Total enrollment`) #create column for graduation rates

```



## Plot graduation numbers and percentages by type of college attended

```{r}
grad_rate_plot <- grad_rate %>%
  ggplot(aes(x = Year,
             y = rate,
             group= College, 
             color= College))+ #plot percent of persons graduating per college type
  geom_line()+ #create line plot
  labs(title = "Number of graduating students by College Type",
       y = "Percent of Graduates")+ #add descriptive titles to figure and legends
  theme_classic()

```

```{r}
college_grad_num <- hbcu_by_school %>%
  ggplot(aes(x = Year,
             y = students,
             group= College, 
             color= College))+ #plot number of persons graduating per college type
  geom_line()+ #create line plot
  labs(title = "Number of graduating students by College Type",
       y = "Graduating Students")+ #add descriptive titles to figure and legends
  theme_classic()

```


# Patch plots together 

```{r}
(grad_rate_plot + college_grad_num) / gender_grad_num  + #Patch all three figure together into 1
   plot_annotation(title = "College Graduation",
                   subtitle = "Fiscal years 1976 - 2015",
                   caption = "Source: Data.World \n 'Bachelor's degree attainment among persons age 25 and over by sex'")+ #add descriptive titles to figure 
  ggsave(here("TidyT_3","outputs", "College_grads.png"), #save to outputs
         width = 10, height = 5)
```

