---
title: "Tidy6"
author: "Anthony Gutierrez"
date: "4/24/2021"
output: html_document
---


# Libraries 

```{r}
library(tidyverse)
library(here)
library(tidytuesdayR) # loading in the TT data
library(cowsay)
```


```{r}
say("woof", by = "cat")
```


# Load Data

```{r}
# This is the transit data
tuesdata6 <- tt_load("2021-01-05")
country_datadic <- read_csv(here("TidyT_6","data", "Country_data_dictionary.csv"))

currency_datadic <- read_csv(here("TidyT_6","data", "Currency_data_dictionary.csv"))

transit <- tuesdata6$transit_cost

```


# Introduction 
Goal of this Tidy Tuesday is to find the average cost of train transit per year by currency type. I will first convert currency to USD. 


# Data Manipulation
I will convert all currency to USD by individually filtering each currency and calculating the conversion. 

############## IS THERE A WAY TO SUMMARIZE THE USD FOR ALL THREE TO CREATE A SINGLE COLUMN ############

# Currency conversion
```{r}
USD_USD <- transit %>%
  mutate(XUSD = case_when(
    currency == "AUD" ~ cost*1.29,
    currency == "BD" ~  cost/84.740,
    currency == "BDT" ~ cost/84.740,
    currency == "BRL" ~ cost/5.5,
    currency == "CAD" ~ cost/1.25,
    currency == "CHF" ~ cost*1.1,
    currency == "CNY" ~ cost/6.25,
    currency == "EGP" ~ cost/15.7,
    currency == "EUR" ~ cost*1.21,
    currency == "INR" ~ cost/74.9,
    currency == "JPY" ~ cost/107.9,
    currency == "USD" ~ cost)) %>%
  filter(complete.cases(.))

view(USD_USD)


#XUSD <- USD_USD %>%
  

```



# Average cost by countries

```{r}

# Box plots of averages by countries that use the listed currencies
USD_USD %>%
  ggplot(aes(y = XUSD, 
             color = currency),#plot cost of transit by currency type
         show.legend = FALSE)+ 
  geom_boxplot()+
  labs(title = "Cost of Train Transportation (USD) per Country currency",
       y = "Cost (USD)")+
  theme_classic()+
  theme(axis.text.x = element_blank())+
  ggsave(here("TidyT_6", "outputs", "Transit_cost.png"), #save to outputs
         width = 7, height = 5) #save figure to outputs
      
```




# Changing prices per year by country
```{r}
USD_USD %>%
  filter(currency == c("CAD", "CNY", "EUR", "INR", "JPY", "USD")) %>% #select for most frequently used currencies in this dataset
  ggplot(aes(x = year,
             y = XUSD, # plot cost of transportation on a yealy basis
             color = country))+ 
  geom_point()+ # points for each counrty that spent x amount of USD
  facet_wrap(~currency, scales = "free")+ #group boxes by currency 
  labs(title = "Cost of Train Transporter by Country Per Year",
       subtitle = "Categorized by Currency",
       x = "Year",
       y = "Cost (USD)",
       legend = "Country")+ #add descriptive titles for all aspects of the figure 
  ggsave(here("TidyT_6", "outputs", "Transit_cost_facet.png"), #save to outputs
         width = 7, height = 5) #save figure to outputs
```








